The Bolocam Galactic Plane Survey: Survey Description and Data Reduction

### 2009-04-17 methods/ms.tex
imagetypes -> negative

% Major Sections
% sec:Introduction
% sec:Observations
% sec:FluxCalibration
% sec:Astrometry
% sec:Mapping
% sec:Noise
% sec:Photometry
% sec:FinalMaps
% sec:Discussion

@@

\author{James Aguirre\altaffilmark{\penn},
        Adam Ginsburg\altaffilmark{\casa},
        Miranda Nordhaus\altaffilmark{\utexas},
	Meredith Drosback\altaffilmark{\virginia},
        John Bally\altaffilmark{\casa}, 
	Cara Battersby\altaffilmark{\casa},
	Eric Todd Bradley\altaffilmark{\ucf},
%        Richard Chamberlin\altaffilmark{\cso},
	Claudia Cyganowski\altaffilmark{\wisc},
	Darren Dowell\altaffilmark{\jpl}
	Neal J. Evans II\altaffilmark{\utexas},
        Jason Glenn\altaffilmark{\casa},
        Paul Harvey\altaffilmark{\utexas,\casa},
        Erik Rosolowsky\altaffilmark{\ubc},
%        Wayne Schlingman\altaffilmark{9},
%        Yancy Shirley\altaffilmark{9},
        Guy S. Stringfellow\altaffilmark{\casa},
        Josh Walawender\altaffilmark{\hawaiihilo}, and 
        Jonathan Williams\altaffilmark{\hawaiimain}
}

@@

This paper details the survey observations and methods.  We treat carefully the
determination of pointing and flux density calibration errors and compare our
results to the literature.  Data processing algorithms that separate
astronomical signals from time-variable atmospheric flux density in the data
time-stream are discussed.  The algorithms reproduce the structure of the
astronomical sky on a limited range of angular scales and produce artifacts in
the vicinity of bright sources. Specifically, extended emission smooth on
scales larger than about 3\arcmin\ is not fully recovered in the images.  The
spatial frequency response function of the Bolocam Galactic Plane Survey is
discussed.

@@

Surveys of the galaxy in CO line emission \citep{dame01,FCRAO} have
identified the locations of large molecular clouds and traced
the transition from atomic to molecular gas as distance from the
center of the galaxy decreases. Together with kinematic distances
\citep[e.g.][]{RotationCurve} and, more recently, accurate distances
from VLBI \citep{reid09}, these surveys have begun to
reveal the spiral structure of the Milky Way. As this picture comes into
sharper focus, we can begin to address the old question of the relationship
between spiral density waves and star formation.

@@

The outline of the paper is as follows. Section \ref{sec:Observations}
describes the instrument and the observations.  Section
\ref{sec:Overview} briefly describes the data reduction process, while
Sections \ref{sec:FluxCalibration} to \ref{sec:FinalMaps} provide
greater detail on flux calibration, astrometry, the mapping algorithm,
noise, photometry, and the public data release.  We conclude in
Section \ref{sec:Discussion} with brief discussion of the immediate
implications of the emission revealed by the BGPS.

@@

\begin{figure}
\hspace{-0.5in}
  \includegraphics[scale=0.8]{fig1_pg4_bw_inv_bar_label} 
  \caption{Cloud complexes centered at $\lon=111$ in the Perseus Arm. \TBD\ Add annotations}
\end{figure}

@@

\begin{figure}
  \hspace{-1in}
  \includegraphics[scale=0.8]{fig1_pg6_bw_inv_bar_label} 
  \caption{W3/4/5 \TBD\ Add annotations}
\end{figure}

@@

% Figure 3
\Figure{bolocam_bandpass}{The Bolocam 1.1 mm bandpass.  Bright
molecular emission lines are shown at their rest frequencies.  Note
that the Bolocam passband rejects $>90\%$ of the $^{12}$CO flux,
leaving our continuum measurements largely uncontaminated.  SO and
CH$_3$OH lines are probably the dominant contributors to line flux in
the passband.  \citet{nummelin1998} found that 22\% of the flux density in one
pointing towards Sgr B2 was from lines, and \citet{yoshida2005}
reported $>40\%$ in Orion A from line emission over the 260-328 GHz band. \TBD\ Other lines
lying in the Bolocam passband include CS($5\to4$) and ($6\to5$) (245
and 293 GHz), HCN($3\to2$) (265 GHz) and HCO$^+(3\to2)$ (267 GHz).}
{fig:Bandpass}{1.0}

@@

This method of alignment avoids the ambiguities inherent in using
extracted sources to align fields, as the Bolocam sources are rarely
point-like.  It further avoids the pointing smearing and loss of peak
flux density which would result if the the maps were coadded using the
pointing model alone, where each individual observation would be
coadded with tthe 6\arcsec\ RMS uncertainty of the model.  The
relative alignment using the correlation maps is typically accurate to
 $<1\arcsec$.

@@

All of the SCUBA 850 and 450 \mum\ data has been re-processed in a
uniform manner and made publicly available as the SCUBA Legacy Catalog
\citep[][hereafter SLC]{difrancesco08}.  These maps provide an excellent cross-check on the
accuracy of the BGPS pointing model.  We have used the same
cross-correlation procedure as that used to obtain the relative
alignment of Bolocam observations to compare to the SLC.  A number of
concerns might arise in comparing the Bolocam to the SCUBA maps,
including the differing scan strategies (SCUBA maps were typically
taken in jiggle-map mode, which tends to remove extended structure)
and the often restricted range of positions measured within an
extended cloud for the SLC.  The cross-correlation technique actually
mitigates these effects, since it does not require the establishment
of a single position for an object (which may differ due to the change
in wavlength between the surveys or due to the observing strategy) but
rather makes use of all the information available.  The
morphological correlation between SCUBA and BGPS sources is generally
excellent, giving a high degree of confidence to the comparison and to
the results obtained from the pointing model.  An example is shown in
Figure \ref{fig:SCUBAPointingComparison}.  We find that the RMS offset
between BGPS and SLC positions within a $3\arcdeg \times 1\arcdeg$
bloc defined by the BGPS is \TBD\ a few arcseconds, consistent with
the errors derived for the Bolocam pointing model.

@@

  \subfigure[{\it Left}: the Bolocam map in a portion of the $\lon=30$
  field.  {\it Right}: the SLC maps in the same region.  The green
  contours on the Bolocam map indicate the positions of the SLC
  sources.  \TBD\ {\it Do the negative bowls of either survey bunge up
  the cross-correlation?  If not, how do we deal with them?  The SCUBA
  image here has been somewhat doctored because of a large DC offset
  in the diagonal filament at center}]
  {\label{fig:SCUBAPointingComparison-a}\includegraphics[scale=0.8]
  {scuba_bolocam_l030}}

@@

We assume the raw timestream data $d$ for each bolometer (indexed by
$i$) can be written as
\begin{equation}
d_i(t) = s_i(t)+a_i(t)+e_i(t)+\epsi_i(t)
\end{equation}
where $s$ is the astrophysical signal, $a$ is atmospheric fluctuation
noise, $e$ are non-random signals due to the instrument itself, and
$\epsi$ is irreducible Gaussian noise due to photon shot and detector
noise.  The process of making a maximum likelihood map in the presence
of Gaussian noise is well-understood, and if the only contributions to
the data were $s$ and $\epsi$, then the minimum variance estimator of
the true astrophysical map is given by
\begin{equation}
\label{eq:LSMap}
m = (A^T W A) A^T W d
\end{equation}
where $A$ encodes the pointing information and $W$ is the inverse of
the covariance matrix of the time domain noise $\epsi$,
\begin{equation}
W = {\langle \epsi_i \epsi_j \rangle}^{-1}
\end{equation}
In this case, the mapping from data to map is a linear operator, $M
=(A^T W A) A^T W$.  For compactness, Equation \ref{eq:LSMap} will be
written as $M[d] = m$.  Note that the mapping operation is not
invertible.  However, given a map $m$ and an observing matrix $A$,
there exists a linear operation which makes predictions about the
observed timestreams, namely $d = A m$; this will be denoted $T[m] =
d$.

@@

We proceed iteratively, estimating $a$ and $e$ in the order of their
relative strengths, beginning with the atmosphere.  The first
iteration step is then
\begin{equation}
\label{eq:FirstIterStep}
\tilde{a}^{(n)}(t) = 
\sum_i{d_i(t)-\tilde{s}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)}
\end{equation}
where $\tilde{s}^{(0)}_i(t)=0$.  This mean atmosphere model is then
fit to each bolometer to obtain the relative gains (flat field) of the
detectors
\begin{equation}
\tilde{a}^{(n)}_i(t) = \frac{d_i(t) \cdot a(t)}{d_i(t)^2} a(t) = r_i a(t)
\end{equation}
This is illustrated in Figure \ref{fig:Flatfield}.  The instrument
errors, which are typically smaller than the atmosphere, are then
estimated
\begin{equation}
\tilde{e}^{(n)}_i(t) = some models
\end{equation}
and both the atmosphere and instrument models are subtracted from the
raw time series
\begin{equation}
d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)
\approx s_i(t) + \epsi_i(t)
\end{equation}
This is the best estimate of signal plus irreducible noise at
iteration step $n$, and is made into a map
\begin{equation}
M[d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)] = \tilde{m}^{(n)}
\end{equation}
The current best map $\tilde{m}^{(n)}$ is then deconvolved to provide
a relatively low-noise, smooth map from which to generate a timestream
\begin{equation}
\label{eq:MapToTime}
T[{\cal D}[\tilde{m}^{(n)}]] = \tilde{s}^{(n)}_i(t)
\end{equation}
where ${\cal D}$ represents the deconvolution operation.  At this
stage, the iterative process could begin again with Equation
\ref{eq:FirstIterStep}, though it is highly useful to produce an
``error map''
\begin{equation}
E = M[d_i(t) - \tilde{s}^{(n)}_i(t) - \tilde{a}^{(n)}_i(t) - 
\tilde{e}^{(n)}_i(t)]
\end{equation}
This error map serves as a straightforward visualization of the
progress of the iterative method, since in a perfect process it would
be map of $\epsi$.

@@

To further remove the atmosphere (and other correlated instrument
noise) without a detailed physical model, we use a principal component
analysis (PCA) as given in \citet{laurent05}.

@@

\item Pickup from the 60 Hz AC power.  This appears as narrow lines in
the PSD of the data.  The second harmonic of 60 Hz is aliased via
beating against the 130 Hz bolometer bias frequency into 10 Hz with
sidebands split at $\sim1$ Hz.

@@

\item Microphonic pickup due to vibrations of the receiver.  The most
noticeable microphonics occur at the end of each scan when the
telescope is turning around and the field rotator is adjusting.  This
leads to broad spikes in the time series, whose long decay must be
removed from the data, particularly during the beginnings of scans.

@@

Fortunately, most of the AC powerline pick-up occurs at frequencies
where there is no astrophysical signal, given the beam size and scan
speed.  Thus this error is dealt with by first notch filtering at the
line frequencies and then low-pass filtering the data.  Because of the
low correlation with astrophysical signal, this step is performed only
once and is not iterated (c.f. Figure XXX of \citet{sayers09}).

Both glitches and microphonic pickup from the scan turnarounds are
degenerate with astrophysical signal and must be estimated as part of
the iterative process.  Glitches are identified as large excursions
from the RMS level after subtraction of the best atmosphere and bright
source model, and the data there is excluded from inclusion in
subsequent maps.  The turnaround microphonics are modeled as decaying
exponentials at the beginnings and ends of scans.

@@

%WRITTEN POORLY: 
The timestream data is made into a spatial map using the pointing data
corresponding to each time point for each bolometer.  The data is
weighted by inverse variance across a single scan and then drizzled
into a map with $7.2$\arcsec\ pixels using a nearest-neighbor
algorithm.  The nearest-neighbor matching allows the map to be
returned to a timestream in the same manner, but with the S/N improved
by averaging over all hits on a given pixel.  It has the disadvantage
that it accentuates the unevenness of the cross-scan sampling, since
ech timestream point's value is assigned to an area much smaller than
the beam.  In principle this can be addressed by various gridding
algorithms, and may be a part of future reductions of the BGPS data.

The resulting map is then subjected to a maximum-entropy CLEAN algorithm
%\citep{hogbom} 
using a specified kernel to produce a deconvolved image
which is the one actually mapped back into the timestream to be
subtracted (Equation \ref{eq:MapToTime}.  This has several advantages
over using the original map directly, including rejecting artefacts
smaller than the beam and decreasing the noise of the resulting
timestream.  It is also better than a ``threshold'' method, which
tends to produce discontinuities.

In some fields, flux from certain source spread out over the course of
the iterative process using certain kernel sizes.  The cause of this
artifact are not yet well-understood, but using a different kernel
size was an effective workaround and did little to change the
properties of the final map.

@@

%\FigureTwo{flagger_withglitch}{flagger_glitchgone}{An illustration of
%the flagging process using the waterfall plot. Left: The glitch
%apparent at middle right is the same one whose time series is shown in
%Figure \ref{fig:CosmicRayGlitch}.  Because of the PCA subtraction, the
%effect of the glitch also propagates over the other bolometers,
%appearing as a horizontal stripe. Right: The same data displayed with
%the time point including the glitch flagged out and rescaled}
%{fig:Flaggin}{1.0}
%
\Figure{crs}{The distribution of glitch amplitudes flagged and removed from the
data in the \lon=111 field.  Glitches below the detection threshold may
contribute to the overall noise signal.}{fig:GlitchDistribution}{1.0}

@@

The error maps produced by the iterative mapping provide a natural way
of estimating the systematic error resulting from imperfect
subtraction of bright sources.  This results in ``ghosts'' of the
bright sources in the error maps $E$.

Noise maps are produced for each map, made from the residuals after
removing the astrophysical signal model as well as the PCA component
model.  These maps are useful for diagnosing the effect of the
iterative process on bright sources and as a method of obtaining an
estimate of the noise.

The noise is correlated between adjacent spatial pixels due to
residual atmospheric noise correlated in time..  We have completed
ignored this effect in our analysis.

The noise may be estimated either from the error maps, from the
timestream PSDs of the detectors, or from the weight maps
produced. The final per-pixel error estimate is given as \TBD.

Because observing conditions varied widely during the survey, the RMS
noise levels obtained in the various fields of the BGPS also varies.
However, the noise within a given $1\arcdeg \times 1\arcdeg$ field is
fairly uniform in the absence of very bright sources.  We show the
variation of the depth as a function of Galactic longitude in Figure
\ref{fig:NoiseVsLongitude}.  The depth can be converted into an
estimate of mass via standard estimates, i.e., under the assumption of
a temperature, opacity, and distance.

Using the same dust opacity as \TBD\ ({\it put in original references})
%\citet{enoch06} and \citet{young06},
and assuming a gas-to-dust mass ratio $X$ of 100, the mass sensitivity
can be written as
\begin{equation} 
\label{eq:Mass}
M_{gas}\approx
0.22\left( e^{12.9/T_d}-1 \right) \left({S_\nu\over 15\; {\rm mJy}} \right)
D^{2}_{kpc} \msol 
\end{equation}
where $D_{kpc}$ is the distance to the source in kpc.  (At $S_\nu =
15$~mJy this corresponds to an RMS extinction sensitivity of $A_V\sim
1$ mag, with the same $T_d$ dependence).  For Bolocam, $h\nu/k \approx
13$~K, so for $T_d > XXX$ the dependence is approximately linear.  At
the coldest temperatures expected, the mass changes a factor of 2.9
between 20 and 10 K.  Far more important is the accurate determination
of the distance to the clump.

@@

To do accurate and meaningful photometry on the Bolocam maps requires
a good estimate of the noise (Section \ref{sec:Noise}) but also an
understanding of the spatial filtering imposed by the observing
strategy and the cleaning and mapping of the data.  Quite apart from
the issues discussed in this section are the issues of identifying
sources with complex morphology and defining the regions over which to
do the aperture.  This is discussed in \citet{rosolowsky09}.  Here we
merely deal with the limitations on performing photometry imposed by
the data analysis itself.

The Bolocam maps produced are the convolution of the (nearly Gaussian)
primary beam $G(\xad)$, together with an effective high-pass
spatial filter $F(\xad)$,
\[
M'(\xad) = (G(\xad) \otimes F(\xad)) \otimes M(\xad)
\]
The effect of the high pass filter is that the effective PSF $B=G
\otimes F$ has the non-intuitive property that its area is {\em zero}:
\[
\int{B(\Omega) d\Omega} = 0
\]
Clearly, this complicates the definition of surface brightness, and
also imposes a limit on the size of features for which photometry may
be meaningfully performed.

@@

The final reduction used 13 PCA components because this produced a
reasonable compromise between attenuating extended structure and
attaining high S/N on compact faint objects.

Figure \ref{fig:deconv_filter}b shows the fraction of flux recovered in the
maps as a function of the deconvolution kernel used to create the astrophysical
model maps.  The differences at large source sizes are not well-understood
(these size scales may be dominated by atmosphere), but at the smallest source
size, it is evident that a deconvolution kernel smaller than the source size
does not reliably recover the source flux.  Therefore we chose a deconvolution
kernel smaller than the beam but larger than the pixel size used.  Figure
\ref{fig:Deconvolution} demonstrates the effects of using both too large and
too small a kernel.

@@

\section{Discussion}
\label{sec:Discussion}

>>>>

% Major Sections
% sec:Introduction
% sec:Observations
% sec:FluxCalibration
% sec:Astrometry
% sec:Mapping
% sec:Noise
% sec:Photometry
% sec:FinalMaps
% sec:Analysis

@@

\author{James Aguirre\altaffilmark{\penn},
        Adam Ginsburg\altaffilmark{\casa},
        Miranda K. Nordhaus\altaffilmark{\utexas},
	Meredith Drosback\altaffilmark{\virginia},
        John Bally\altaffilmark{\casa}, 
	Cara Battersby\altaffilmark{\casa},
	Eric Todd Bradley\altaffilmark{\ucf},
%        Richard Chamberlin\altaffilmark{\cso},
	Claudia Cyganowski\altaffilmark{\wisc},
	Darren Dowell\altaffilmark{\jpl}
	Neal J. Evans II\altaffilmark{\utexas},
        Jason Glenn\altaffilmark{\casa},
        Paul Harvey\altaffilmark{\utexas,\casa},
        Erik Rosolowsky\altaffilmark{\ubc},
%        Wayne Schlingman\altaffilmark{9},
%        Yancy Shirley\altaffilmark{9},
        Guy S. Stringfellow\altaffilmark{\casa},
        Josh Walawender\altaffilmark{\hawaiihilo}, and 
        Jonathan Williams\altaffilmark{\hawaiimain}
}

@@

This paper details the survey observations and methods.  We treat carefully the
determination of pointing and flux density calibration errors and compare our
results to the literature.  Data processing algorithms that separate
astronomical signals from time-variable atmospheric flux density in the data
time-stream are discussed.  The algorithms reproduce the structure of the
astronomical sky on a \emph{limited range of angular scales} and produce artifacts in
the vicinity of bright sources. Specifically, extended emission smooth on
scales larger than about 3\arcmin\ is not fully recovered in the images.  The
spatial frequency response function of the Bolocam Galactic Plane Survey is
discussed.

@@

Surveys of the galaxy in CO line emission \citep{dame01,jackson06} have
identified the locations of large molecular clouds and traced
the transition from atomic to molecular gas as distance from the
center of the galaxy decreases. Together with kinematic distances
\citep[e.g.][]{RotationCurve} and, more recently, accurate distances
from VLBI \citep{reid09}, these surveys have begun to
reveal the spiral structure of the Milky Way. As this picture comes into
sharper focus, we can begin to address the old question of the relationship
between spiral density waves and star formation.

@@

The outline of the paper is as follows. Section \ref{sec:Observations}
describes the instrument and the observations.  Section
\ref{sec:Overview} briefly describes the data reduction process, while
Sections \ref{sec:FluxCalibration} to \ref{sec:FinalMaps} provide
greater detail on flux calibration, astrometry, the mapping algorithm,
noise, photometry, and the public data release.  We conclude in
Section \ref{sec:Analysis} with brief discussion of the immediate
implications of the emission revealed by the BGPS.

@@

\begin{figure}
\hspace{-0.5in}
  \includegraphics[scale=0.8]{fig1_pg4_bw_inv_bar_label} 
  \caption{Cloud complexes centered at $\lon=111$ in the Perseus Arm. 
  The NGC 7538 complex is in the upper-left.}
\end{figure}

@@

\begin{figure}
  \hspace{-1in}
  \includegraphics[scale=0.8]{fig1_pg6_bw_inv_bar_label} 
  \caption{W3/4/5.  The W3(OH)/W3 Main complex is the bright source
  on the right side of the image.  W5 has one scan performed in RA/Dec
  instead of Galactic coordinates and so has non-uniform noise properties
  in each square degree}
\end{figure}

@@

% Figure 3
\Figure{bolocam_bandpass}{The Bolocam 1.1 mm bandpass.  Bright molecular
emission lines are shown at their rest frequencies.  Note that the Bolocam
passband rejects $>90\%$ of the $^{12}$CO flux, leaving our continuum
measurements largely uncontaminated.  SO and CH$_3$OH lines are probably the
dominant contributors to line flux in the passband.  \citet{nummelin1998} found
that 22\% of the flux density in one pointing towards Sgr B2 was from lines,
and \citet{yoshida2005} reported $>40\%$ in Orion A from line emission over the
260-328 GHz band. Other lines lying in the Bolocam passband include CS($5\to4$)
and ($6\to5$) (245 and 293 GHz), HCN($3\to2$) (265 GHz) and HCO$^+(3\to2)$ (267
GHz).  A follow-up project will attempt to determine the contribution of lines
to the BGPS flux density measurements.}
{fig:Bandpass}{1.0}

@@

This method of alignment avoids the ambiguities inherent in using
extracted sources to align fields, as the Bolocam sources are rarely
point-like.  It further avoids the pointing smearing and loss of peak
flux density which would result if the the maps were coadded using the
pointing model alone, where each individual observation would be
coadded with the 6\arcsec\ RMS uncertainty of the model.  The
relative alignment using the correlation maps is typically accurate to
 $<1\arcsec$.

@@

All of the SCUBA 850 and 450 \mum\ data has been re-processed in a uniform
manner and made publicly available as the SCUBA Legacy Catalog
\citep[][hereafter SLC]{difrancesco08}.  These maps provide an excellent
cross-check on the accuracy of the BGPS pointing model.  We have used the same
cross-correlation procedure as that used to obtain the relative alignment of
Bolocam observations to compare to the SLC.  A number of concerns might arise
in comparing the Bolocam to the SCUBA maps, including the differing scan
strategies (SCUBA maps were typically taken in jiggle-map mode, which tends to
remove extended structure) and the often restricted range of positions measured
within an extended cloud for the SLC.  The cross-correlation technique actually
mitigates these effects, since it does not require the establishment of a
single position for an object (which may differ due to the change in wavlength
between the surveys or due to the observing strategy) but rather makes use of
all the information available.  The morphological correlation between SCUBA and
BGPS sources is generally excellent with some exceptions, giving a high degree
of confidence to the comparison and to the results obtained from the pointing
model.  An example is shown in Figure \ref{fig:SCUBAPointingComparison}.  We
find that the RMS offset between BGPS and SLC positions within a $3\arcdeg
\times 1\arcdeg$ block defined by the BGPS is $4.0\pm8.8$\arcsec, consistent
with the errors derived for the Bolocam pointing model.  There were some BGPS
fields that had larger offsets from SCUBA fields; these fields were excluded
from the offset measurement above because of poor morpohological match.

@@

  \subfigure[{\it Left}: the Bolocam map in a portion of the $\lon=30$
  field.  {\it Right}: the SLC maps in the same region.  The green
  contours on the Bolocam map indicate the positions of the SLC
  sources.  Negative bowls in the image are forced to zero before
  cross-correlating.  ]
  {\label{fig:SCUBAPointingComparison-a}\includegraphics[scale=0.8]
  {scuba_bolocam_l030}}

@@

We assume the raw timestream data $d$ for each bolometer (indexed by
$i$) can be written as
\begin{equation}
d_i(t) = s_i(t)+a_i(t)+e_i(t)+\epsi_i(t)
\end{equation}
where $s$ is the astrophysical signal, $a$ is atmospheric fluctuation
noise, $e$ are non-random signals due to the instrument itself, and
$\epsi$ is irreducible Gaussian noise due to photon shot and detector
noise.  The process of making a maximum likelihood map in the presence
of Gaussian noise is well-understood, and if the only contributions to
the data were $s$ and $\epsi$, then the minimum variance estimator of
the true astrophysical map is given by
\begin{equation}
\label{eq:LSMap}
m = (A^T W A) A^T W d
\end{equation}
where $A$ encodes the pointing information and $W$ is the inverse of
the covariance matrix of the time domain noise $\epsi$,
\begin{equation}
W = {\langle \epsi_i \epsi_j \rangle}^{-1}
\end{equation}
In this case, the mapping from data to map is a linear operator, $M
=(A^T W A) A^T W$.  For compactness, Equation \ref{eq:LSMap} will be
written as $M[d] = m$.  Note that the mapping operation is not
invertible.  However, given a map $m$ and an observing matrix $A$,
there exists a linear operation which makes predictions about the
observed timestreams, namely $d_p = A m$; this will be denoted $T[m] =
d_p$.

@@

We proceed iteratively, estimating $a$ and $e$ in the order of their
relative strengths, beginning with the atmosphere.  The first
iteration step is then
\begin{equation}
\label{eq:FirstIterStep}
\tilde{a}^{(n)}(t) = 
\sum_i{d_i(t)-\tilde{s}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)}
\end{equation}
where $\tilde{s}^{(0)}_i(t)=0$.   Each bolometer's timestream is then 
fit to the mean atmosphere model to obtain the relative gains of the detectors.
\begin{equation}
\tilde{a}^{(n)}_i(t) = \frac{d_i(t) \cdot a(t)}{d_i(t)^2} a(t) = r_i a(t)
\end{equation}
This is illustrated in Figure \ref{fig:Flatfield}.  
% James - what's this section about?
% The instrument
% errors, which are typically smaller than the atmosphere, are then
% estimated
% \begin{equation}
% \tilde{e}^{(n)}_i(t) = some models
% \end{equation}
Both the atmosphere and instrument models are subtracted from the
raw time series
\begin{equation}
d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)
\approx s_i(t) + \epsi_i(t)
\end{equation}
This is the best estimate of signal plus irreducible noise at
iteration step $n$, and is made into a map
\begin{equation}
M[d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)] = \tilde{m}^{(n)}
\end{equation}
The current best map $\tilde{m}^{(n)}$ is then deconvolved to provide
a relatively low-noise, smooth map from which to generate a timestream
\begin{equation}
\label{eq:MapToTime}
T[{\cal D}[\tilde{m}^{(n)}]] = \tilde{s}^{(n)}_i(t)
\end{equation}
where ${\cal D}$ represents the deconvolution operation.  At this
stage, the iterative process could begin again with Equation
\ref{eq:FirstIterStep}, though it is useful to produce a
``residual map''
\begin{equation}
E = M[d_i(t) - \tilde{s}^{(n)}_i(t) - \tilde{a}^{(n)}_i(t) - 
\tilde{e}^{(n)}_i(t)]
\end{equation}
This residual map serves as a visualization of the progress of the iterative
method, since in a perfect process it would be map of $\epsi$.

@@

To further remove the atmosphere (and other correlated instrument
noise) without a detailed physical model, we use principal component
analysis (PCA) as given in \citet{laurent05}.

@@

\item Pickup from the 60 Hz AC power.  This appears as narrow lines in the PSD
    of the data.  The second harmonic of 60 Hz is aliased via beating against
    the 130 Hz bolometer bias frequency into 10 Hz with sidebands split at
    $\sim1$ Hz.  It is removed by replacing components at these frequencies in
    the fourier transform with gaussian noise matched to the local mean
    amplitude.

@@

% James - technical question: Why is there microphonic pickup?  What
% does the term mean, exactly?
\item Microphonic pickup due to vibrations of the receiver.  The most
noticeable microphonics occur at the end of each scan when the
telescope is turning around and the field rotator is adjusting.  This
leads to broad spikes in the time series, whose long decay must be
removed from the data, particularly during the beginnings of scans.

@@

Fortunately, most of the AC powerline pick-up occurs at frequencies
where there is no astrophysical signal, given the beam size and scan
speed.  Thus this error is dealt with by first notch filtering at the
line frequencies and then low-pass filtering the data.  Because of the
low correlation with astrophysical signal, this step is performed only
once and is not iterated (c.f. Figure 1 of \citet{sayers09}).

Both glitches and microphonic pickup from the scan turnarounds are
degenerate with astrophysical signal and must be estimated as part of
the iterative process.  
% NOT TRUE - they are identified per-pixel most of the time
% Glitches are identified as large excursions
% from the RMS level after subtraction of the best atmosphere and bright
% source model, and the data there is excluded from inclusion in
% subsequent maps.
The turnaround microphonics are modeled as decaying exponentials at the
beginnings and ends of scans and are also removed by the PCA cleaning.

@@

The timestream data is made into a spatial map using the pointing data
corresponding to each time point for each bolometer.  The data is
weighted by inverse variance across a single scan and then drizzled
into a map with $7.2$\arcsec\ pixels using a nearest-neighbor
algorithm.  The nearest-neighbor matching allows the map to be
returned to a timestream in the same manner, but with the S/N improved
by averaging over all hits on a given pixel.  It has the disadvantage
that it accentuates the unevenness of the cross-scan sampling, since
each timestream point's value is assigned to an area much smaller than
the beam.
%In principle this can be addressed by various gridding
%algorithms, and may be a part of future reductions of the BGPS data.

The resulting map is then subjected to a maximum-entropy clean algorithm
\citep{hollis92} using a specified kernel to produce a deconvolved image which
is mapped back into the timestream to be subtracted (Equation
\ref{eq:MapToTime}).  This has several advantages over using the original map
directly, including rejecting artifacts smaller than the kernel and decreasing
the noise of the resulting timestream.  It is also better than a ``threshold''
method, which tends to produce discontinuities and include noise outliers.

In some fields, flux from some sources spread out over the course of the
iterative process using certain kernel sizes.  The cause of this artifact is
not yet well-understood, but using a different kernel size was an effective
workaround and did little to change the properties of the final map.  The
standard kernel size was 14.4\arcsec; a 21.6\arcsec kernel was used when
artifacts were encountered.

@@

%\FigureTwo{flagger_withglitch}{flagger_glitchgone}{An illustration of
%the flagging process using the waterfall plot. Left: The glitch
%apparent at middle right is the same one whose time series is shown in
%Figure \ref{fig:CosmicRayGlitch}.  Because of the PCA subtraction, the
%effect of the glitch also propagates over the other bolometers,
%appearing as a horizontal stripe. Right: The same data displayed with
%the time point including the glitch flagged out and rescaled}
%{fig:Flaggin}{1.0}
%
\Figure{crs}{The distribution of glitch amplitudes flagged and removed from the
data in the \lon=111 field.  Glitches below the detection threshold may
contribute to the overall noise.}{fig:GlitchDistribution}{1.0}

@@

The residual maps produced by the iterative mapping provide a natural way of
estimating the systematic error resulting from imperfect subtraction of bright
sources.  The imperfect subtraction results in ``ghosts'' of the bright sources
in the error maps $E$.

Residual maps are produced for each map from the residuals after
removing the astrophysical signal model and the PCA atmosphere model.  These
maps are useful for diagnosing the effect of the iterative process on bright
sources and as a method of obtaining an
estimate of the noise.

% ...ok, explain please?
% The noise is correlated between adjacent spatial pixels due to
% residual atmospheric noise correlated in time.  We have completed
% ignored this effect in our analysis.

The noise may be estimated either from the residual maps, from the timestream
PSDs of the detectors, or from the weight maps produced. The final per-pixel
error estimate is given as \TBD (this remains an open question - is the best
per-pixel error a smoothed version of the residual map, e.g.  \citet{enoch06},
1/weightmap$^2$, or some other estimate? Adam, James, Erik).

Because observing conditions varied widely during the survey, the RMS
noise levels obtained in the various fields of the BGPS also varies.
However, the noise within a given $1\arcdeg \times 1\arcdeg$ field is
fairly uniform in the absence of very bright sources.  We show the
variation of the depth as a function of Galactic longitude in Figure
\ref{fig:NoiseVsLongitude}.

The depth can be converted into an
estimate of mass via standard estimates, i.e., under the assumption of
a temperature, opacity, and distance.
Using dust opacity from \citet{ossenkopf94}
%\citet{enoch06} and \citet{young06},
and assuming a gas-to-dust mass ratio 100, the mass sensitivity
can be written as
\begin{equation} 
\label{eq:Mass}
M_{gas}\approx
0.22\left( e^{13.1/T_d}-1 \right) \left({S_\nu\over 15\; {\rm mJy}} \right)
D^{2}_{kpc} \msol 
\end{equation}
where $D_{kpc}$ is the distance to the source in kpc.  (At $S_\nu =
15$~mJy this corresponds to an RMS extinction sensitivity of $A_V\sim
1$ mag, with the same $T_d$ dependence).  For Bolocam, $h\nu/k \approx
13$~K, so for $T_d > 10$K the dependence is approximately linear.  At
the coldest temperatures expected, the mass changes a factor of 2.9
between 20 and 10 K.  The distance to the clump is more important in most
situations.

@@

To do accurate and meaningful photometry on the Bolocam maps requires
a good estimate of the noise (Section \ref{sec:Noise}) but also an
understanding of the spatial filtering imposed by the observing
strategy and the cleaning and mapping of the data.  Quite apart from
the issues discussed in this section are the issues of identifying
sources with complex morphology and defining the regions over which to
do the aperture.  This is discussed in \citet{rosolowsky09}.  Here we
merely deal with the limitations on performing photometry imposed by
the data itself.

The Bolocam maps produced are the convolution of the (nearly Gaussian)
primary beam $G(\xad)$, together with an effective high-pass
spatial filter $F(\xad)$ (figure \ref{fig:PCA_filter}),
\[
M'(\xad) = (G(\xad) \otimes F(\xad)) \otimes M(\xad)
\]

%James - please explain this to me, or explain it better here.  What
%is the form of B(Omega) that it gives 0?
The effect of the high pass filter is that the effective PSF $B=G
\otimes F$ has the non-intuitive property that its area is {\em zero}:
\[
\int{B(\Omega) d\Omega} = 0
\]
Clearly, this complicates the definition of surface brightness, and
also imposes a limit on the size of features for which photometry may
be meaningfully performed.

@@

The final reduction used 13 PCA components because this produced a
reasonable compromise between attenuating extended structure and
cleaning the atmospheric contribution.  Note that this is a substantially
more aggressive cleaning than that used in \citet{enoch06} largely because
our mean integration time per pixel was $\sim$seconds, whereas theirs was
$\sim$minutes, and therefore the atmospheric noise was much more likely to
be correlated between observations.

Figure \ref{fig:PCA_Filter}b shows the fraction of flux recovered in the
maps as a function of the deconvolution kernel used to create the astrophysical
model maps.  The differences at large source sizes are not well-understood
(these size scales may be dominated by atmosphere), but at the smallest source
size, it is evident that a deconvolution kernel smaller than the source size
does not reliably recover the source flux.  Therefore we chose a deconvolution
kernel smaller than the beam but larger than the pixel size used.  Figure
\ref{fig:Deconvolution} demonstrates the effects of using both too large and
too small a kernel.

@@

\section{Analysis}
\label{sec:Analysis}

The BGPS is the first millimeter survey of a substantial fraction
of the Galactic Plane, providing an unbiased look at the high-density
gas most intimately associated with the earliest phases of star
formation.
